First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-03 referee_response.txt

Major comments
---------------
---The authors emphasize the diagnostic importance of the kperp-kpara plot, and I appreciate the care with which such plots were discussed. However, I found the absence of a kpara-kperp error plot to be rather conspicuous. I would imagine that an error plot would also be enormously informative, and I'd like to see one included, along with a short discussion of its features.

We have added an error plot in Section 3.3 (now Figure 3), along with a paragraph at the end of the section describing the structure.

======================================================================

---11th paragraph of Section 3.1. "At this stage we also average in frequency by a factor of two..." Is there a specific motivation for this? Or is it just a convenient way to reduce the data volume?

======================================================================

---After Eq. (5): "After fitting for all other parameters, we fit for the reflection mode". Why do this instead of a global fit for all the parameters, which would presumably give a better overall fit? I've noticed that with the form of the g-hat, if A_i,1 and A_i,2 are zero, the reflection mode parameters are degenerate with the "normal" calibration parameters. Is the separate fit a desire to avoid the near-degeneracy that results when A_i,1 and A_i,2 are small?

======================================================================

---Penultimate paragraph of Section 4.2. This reminds me of the Hazelton et al. (2013) paper about multi-baseline wedge effects. It would seem like the current paper is suggesting that the Hazelton et al. effect can be reduced with a fine kernel resolution. However, Hazelton et al. seem to imply that their effect is fundamental and unavoidable. Can the authors explicitly clarify this?

======================================================================

---2nd paragraph of Section 4.3. "All other sources were given a spectral index of -0.8". What motivated this choice?

======================================================================

---Penultimate paragraph of Section 4.3. The authors find that using the two different foreground models gave different flux scales. Is there a sense in which the flux scale has converged?

======================================================================

---Figures 9 and 10. Figure 10 seems to indicate that there is a "+5" pointing. Why is it not included in Figure 9?

======================================================================

---Figure 11, bottom panel. Since the "+3" portion seems so tightly clustered and is only excluded from the analysis due to some consistent offset, might this be fixable by adjusting calibration parameters?

======================================================================

---8th paragraph of Section 5.2. I think the vertical lines caused by incomplete uv coverage needs more explanation/discussion. A thinning uv coverage alone should not result in the vertical streaks. It needs to be coupled with increased leakage from within the wedge. The fact that reduced uv coverage causes vertical streaks is therefore also a result of how the power spectrum is estimated (since that + the instrument is what determines the leakage). As an example, one could go with a "scorched earth" style analysis, sacrificing sensitivity to preserve the EoR window by doing single-baseline delay spectrum analyses for the longer baselines. The foregrounds would be contained within the wedge under such a procedure, independent of uv-coverage (since we're already at the extreme of analyzing one baseline at a time). The form of one's power spectrum estimator must therefore play a role here.

======================================================================

---Along the same lines as the previous point, might an accounting of covariances between different power spectrum cells also help here, perhaps allowing for decreased leakage?

======================================================================

---5th from last paragraph of Section 5.2. Why an increase in slope of 14% of the horizon line? I'm curious as to the thinking here, particularly versus the alternative I've seen in the literature where a constant delay buffer is added to the horizon line.

======================================================================

---Figure 15. Why does the FHD + eppsilon pipeline not go to as high k as RTS + CHIPS? Is there some limitation that prevents the highest k regions from being explored?

======================================================================

---It would be helpful in the discussion section to give a sense for what data and results we can look forward to. As an outsider, for example, I am curious as to whether there is more data "in the can" that is just waiting to be analyzed, or whether any new results will come from data taken once the MWA expansion is past its construction/commissioning phase.

======================================================================

Minor comments
---------------
---Throughout: the authors seem to use \citealt{} or \citep{} for all citations, but when the citation is part of the sentence, \citet{} to be used (so that publication year but not the author name is enclosed in parentheses).

This has been fixed in the updated draft.

======================================================================

---4th paragraph of introduction: Greig et al. also placed constraints on physical models, I think.

Yes, thank you for pointing that out. This citation has been added.

======================================================================

---Final sentence of 4th paragraph of introduction: The two halves of the sentence don't really have a logical connection to each other. I was confused for a moment because I thought there was a missing part to sentence. In particular, when I read "...and several analysis pipelines are under active development" I thought the authors meant that there were specific efforts to preserve the EoR window, but I think it's just generic pipeline development.

We have rephrased to “… and several analysis pipelines are under active development to exploit this foreground isolation” to emphasize that the pipelines cited are following the EoR window framework.

======================================================================

---Similarly, in the last paragraph of Section 2, when the authors say "(e.g., Jacobs et al. 2016...)", I'm not sure what the "for example" is referring to. I'm not sure how those other works demonstrated that "all techniques demonstrated here will use this golden data set"?

Those citations were misplaced and meant for the previous sentence. This has been fixed.

======================================================================

---Equation 1: V_ij is never defined as the "true" visibility.

It is now defined in the sentence after the equation.

======================================================================

---11th paragraph of Section 3.1 and first paragraph of Section 3.3: what does "frame" mean? Is this a technical term? I'm not familiar with it.

“Holographic frame” is now introduced as a term referring to the specific weighting scheme that yields optimal maps. We also removed another use of “frame” in Sec 4.4 when referring to weighting.

======================================================================

---Section 3.3: even though it can be inferred from the description in Section 3.2, it would be helpful to readers to mention the foreground subtraction step in Section 3.3 to make it clear where it fits into the power spectrum pipeline.

We have injected a paragraph to Section 3.3 to mention the foreground subtraction. This is now the third paragraph of the section.

======================================================================

---Section 3.3, 3rd paragraph: "The three dimensional power spectrum cube can next be averaged in annuli to form..." For readers who may not be accustomed to P(kperp, kpara) spectra, it would be good to be more specific here. (E.g., annuli oriented in which direction?)

This is now clarified as the annuli of constant (kperp, kpar), (i.e. orthogonal to the kpar axis).

======================================================================

---In all kperp-kpara plots: The units "ns" and "lambda" are given, but it's never stated what quantities (i.e., "delay" and "baseline length", respectively) are plotted.

We have added a description of these axes in the penultimate paragraph of Section 3.3 (and in the caption of Figure 2).

======================================================================

---I don't think "KGS" is ever defined.

It is now spelled out in a footnote when we introduce the catalog.

======================================================================

---Just before Equation 6. "Pixel" is a bit of a confusing term here. Perhaps something like "kperp-kpara cell" might be more appropriate?

This has been changed.

======================================================================

---An inconsistency in citation style. Why is "J. L. B. Line" always given his initials while everyone else is cited by last name only?

Our understanding is that unpublished references should include the initials, while published references do not (http://journals.aas.org/authors/references.html#Unpublished). However, when double checking we noticed a couple unpublished references that were missing the initials, and have now added them.

======================================================================

---Penultimate paragraph of Section 4.4. What does "total residual power" mean? Can we be more precise? Is it the power integrated over the whole kperp-kpara plane? If so, the relevant k scales of integration need to be specified.

======================================================================

---Last paragraph of Section 4.4. I am having trouble understanding the sentence "The top panel shows the model power with the diffuse subtracted from the model power without". Can we rephrase? Also, I'm a little confused by what "model power with the diffuse subtracted" means. Isn't the diffuse power "added" to the model on top of the point sources?

This was worded in a very confusing way. The sentence now reads: “The top panel shows our point source foreground model power spectrum minus the model power spectrum when including diffuse.”

======================================================================

---Figure 8 caption: I don't understand the meaning of the parenthetical remark "(positive difference)".

This sentence had a lot of words to orient the reader between the color scheme corresponding to positive/negative, and which power spectrum was subtracted from which. We have broken it up into two sentences to be more clear.

======================================================================

---Figure 8 caption (but also in the text): is it not a bit optimistic to say that the diffuse modeling "shows promise" just because 90% of power is removed? 90% sounds impressive if we think in linear terms, but on a logarithmic scale (which is what counts), it's just one order of magnitude out of many.

We have removed the subjective language.

======================================================================

---Final paragraph of the paper. It's not quite true that imaging is necessary for cross-correlation. One could also take the images from other surveys and use them to predict visibilities, allowing for a cross-correlation in visibility space.

While this is technically true (and interesting), it does not avoid the challenges inherent in imaging. In order to achieve the visibility correlation, one would need to simulate the measured visibilities with their phase to high accuracy. In the literature we often lump all instrumental effects into the "A" matrix operating on a sky vector resulting in a set of measurements, assuming the calibration will fix all the small instrumental effects and make the real A match the model A. For a cross-correlation one would need to simulate with the real A, including phase corrections at the level needed for imaging. This can be sidestepped in power spectrum measurements because localization is irrelevant, but it is essential for cross-correlation, and the calibration uncertainties leaks back in in the difference between simulating with a model A and the real A. For all intents and purposes, cross-correlating in visibility space would still be an imaging analysis. There is actually a paper in the works (M.F. Morales et al) which will very carefully explore the definition and trade-offs of imaging analysis.

### 2016-08-04 referee_response.txt

We have added an error plot in Section 3.3 (now Figure 3), along with a paragraph at the end of the section describing the structure.

======================================================================

---11th paragraph of Section 3.1. "At this stage we also average in frequency by a factor of two..." Is there a specific motivation for this? Or is it just a convenient way to reduce the data volume?

Yes, this is done to reduce data volume. We have added a sentence explaining this, and pointing out that it does not affect our sensitivity.

@@

======================================================================

---After Eq. (5): "After fitting for all other parameters, we fit for the reflection mode". Why do this instead of a global fit for all the parameters, which would presumably give a better overall fit? I've noticed that with the form of the g-hat, if A_i,1 and A_i,2 are zero, the reflection mode parameters are degenerate with the "normal" calibration parameters. Is the separate fit a desire to avoid the near-degeneracy that results when A_i,1 and A_i,2 are small?

That’s essentially it. We found that if we don’t first subtract the polynomial fit, our reflection fits were dominated by the low order terms, and the results did not make physical sense. Even though the degeneracy should be low (the reflection ripple is quite fine compared to other parameters), the relatively huge amount of power in the polynomial terms was completely obscuring the reflection fits. We have added a sentence to this effect in the text.

@@

======================================================================

---Penultimate paragraph of Section 4.2. This reminds me of the Hazelton et al. (2013) paper about multi-baseline wedge effects. It would seem like the current paper is suggesting that the Hazelton et al. effect can be reduced with a fine kernel resolution. However, Hazelton et al. seem to imply that their effect is fundamental and unavoidable. Can the authors explicitly clarify this?

This effect is distinct from other wedge features discussed in the literature, and we have added a paragraph (3rd paragraph of Sec 4.2) pointing this out. The main difference is that this effect is due to the computational limits of the analysis itself, whereas other effects are inherent in the data. In the specific case of Hazelton et al. (2013), the effect is due to multiple baselines migrating in and out of uv cells, and the effective weighting of a given cell is frequency dependent. While the exact mixing would be dependent on the gridding kernel used, it would still exist in the limit of infinitely precise models because the frequency dependent weighting is due to the inherent chromaticity of the instrument.

@@

---5th from last paragraph of Section 5.2. Why an increase in slope of 14% of the horizon line? I'm curious as to the thinking here, particularly versus the alternative I've seen in the literature where a constant delay buffer is added to the horizon line.

======================================================================

---Figure 15. Why does the FHD + eppsilon pipeline not go to as high k as RTS + CHIPS? Is there some limitation that prevents the highest k regions from being explored?

This is simply due to the frequency averaging FHD+eppsilon does at the gridding step, which RTS+CHIPS does not do. This is now explained in the caption of Fig. 15.

### 2016-08-04 ms.tex
Addressing more referee major comments

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the \textbf{``holographic frame,"} which has been shown to be the optimal \textbf{weighting scheme} for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

\textbf{Figure~\ref{fig:example_2derror} shows the error level corresponding to the two 
dimensional power spectrum in Figure~\ref{fig:example_ps}. These errors
are found by propagating noise measured from calibrated visibilities through the entire
pipeline. The dense core of the MWA results in relatively low noise at low $k_\perp$.
As the $(u,v)$ density of the array decreases at longer baselines, the noise increases.
The noise is mostly constant across $k_{||}$ because the Fourier transform distributes
the noise in frequency to all modes equally \citep{Morales:2004}.
The exception is the second $k_{||}$ bin, which has slightly lower noise than the rest of the 
$k_{||}$ modes. This is a result of the Lomb-Scargle periodogram interacting with the
Blackman-Harris window function we apply to the data. In the case of $k_{||}=0$
(shown as the bottom bin on the log plot), the
cosine term carries all the information and has a noise level $\sqrt{2}$ lower than any other
sine or cosine term individually. The first non-zero $k_{||}$ mode is highly covariant with the 
bottom bin due to the window function. Therefore the noise on this mode's cosine term is 
lower than other higher order cosines. With the added contribution from the sine term, this
mode contains the lowest level of noise.
}

@@

The top panel of Figure~\ref{fig:cables} shows a 2D power spectrum using the gain model 
described so far. Three of the horizontal lines in the EoR window can be attributed to the 
coarse band gaps, as they reside at the harmonics expected for 1.28~MHz periodic 
sampling. But the sampling cannot account for the fourth line, highlighted by the arrow. The 
corresponding delay time of the line, $\tau \approx 1.23$~$\mu$s, corresponds almost 
exactly to twice the signal travel time through the 150 meter cables (with velocity 0.81c). We 
therefore introduced a reflection term into our gain model, which allows our restricted gains 
to account for the interference of the incident signal with a round-trip reflected signal in the 
cable. 
A similar approach was taken at lower frequencies in \citet{Ewall-Wice:2016}.
The full gain model is expressed as
\begin{equation}\label{eq:cal2}
\hat{g}_i(\nu)=B_{\alpha}(\nu)\left(P_i(\nu)+R_i(\nu)\right),
\end{equation}
where the reflection term can be further decomposed as
\begin{equation}
R_i(\nu) = R_{i,0} e^{-2\pi i \tau_{i} \nu}.
\end{equation}
The reflection coefficient, $R_{i,0}$ is allowed to be complex, while the reflection delay, 
$\tau_i$ is real. After fitting for all other parameters, we fit for the reflection mode. This is 
done for antennas with suspected reflections, the most offensive of which is seen in the 
antennas with cables of 150 meters. Because the cables were not cut at exact lengths 
(variations on order tens of centimeters), we solve for both the reflection coefficient and 
delay by performing a direct Fourier transform to a highly over resolved delay grid and 
selecting the mode where the reflection amplitude is largest. 
This fitting produces a single reflection coefficient for each antenna with nominal cable
length of 150 meters.

@@

@@

\begin{figure}
\includegraphics[width=\columnwidth]{1D_spectra_comparison.pdf}
\caption{
Here we compare the upper limits from the FHD$+$\eppsilon pipeline with those from the 
RTS$+$CHIPS pipeline. The magenta lines are repeated from Figure~\ref{fig:1d_ps}, and the gray lines
are the power plus two sigma thermal error bars lines generated by CHIPS. 
Due to the logarithmic binning used by CHIPS, some bins are empty, resulting in gaps in the
spectra.
CHIPS does not 
excise the coarse band harmonic lines, and so we see their strong effect where there are 
gaps in the magenta lines. The most notable differences are the ability of the RTS$+$CHIPS pipe
to access lower $k$ modes, while the FHD$+$\eppsilon pipe suppresses more power before and
after the first coarse band harmonic. While the different analysis methods result in slightly
different upper limits, they are generally in agreement. 
\label{fig:ps_comparison}
}
\end{figure}

>>>>

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the \textbf{``holographic frame,"} which has been shown to be the optimal \textbf{weighting scheme} for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz. 
\textbf{This is done to reduce the data volume, and has little impact on our results because
our sensitivity to the EoR is extremely low at the corresponding line-of-sight $k$ modes
\citep{Beardsley:2013}}

@@

\textbf{Figure~\ref{fig:example_2derror} shows the error level corresponding to the two 
dimensional power spectrum in Figure~\ref{fig:example_ps}. These errors
are found by propagating noise measured from calibrated visibilities through the entire
pipeline. The dense core of the MWA results in relatively low noise at low $k_\perp$.
As the $(u,v)$ density of the array decreases at longer baselines, the noise increases.
The noise is mostly constant across $k_{||}$ because the Fourier transform distributes
the noise in frequency to all modes equally \citep{Morales:2004}.
The exception is the second $k_{||}$ bin, which has slightly lower noise than the rest of the 
$k_{||}$ modes. 
As explained in Hazelton et al. (in prep.) this is a result of the Lomb-Scargle periodogram 
interacting with the Blackman-Harris window function we apply to the data. Most of the bins 
have the noise equally distributed between cosine and sine terms, with the exception of the 
$k_{||}=0$ bin where the cosine term carries all the information. Due to the correlations 
introduced by the Blackman-Harris window function the first non-zero $k_{||}$ mode is highly 
covariant with the bottom bin ? therefore the noise on this mode's cosine term is lower than 
other higher order cosines.
}

@@

The top panel of Figure~\ref{fig:cables} shows a 2D power spectrum using the gain model 
described so far. Three of the horizontal lines in the EoR window can be attributed to the 
coarse band gaps, as they reside at the harmonics expected for 1.28~MHz periodic 
sampling. But the sampling cannot account for the fourth line, highlighted by the arrow. The 
corresponding delay time of the line, $\tau \approx 1.23$~$\mu$s, corresponds almost 
exactly to twice the signal travel time through the 150 meter cables (with velocity 0.81c). We 
therefore introduced a reflection term into our gain model, which allows our restricted gains 
to account for the interference of the incident signal with a round-trip reflected signal in the 
cable. 
A similar approach was taken at lower frequencies in \citet{Ewall-Wice:2016}.
The full gain model is expressed as
\begin{equation}\label{eq:cal2}
\hat{g}_i(\nu)=B_{\alpha}(\nu)\left(P_i(\nu)+R_i(\nu)\right),
\end{equation}
where the reflection term can be further decomposed as
\begin{equation}
R_i(\nu) = R_{i,0} e^{-2\pi i \tau_{i} \nu}.
\end{equation}
The reflection coefficient, $R_{i,0}$ is allowed to be complex, while the reflection delay, 
$\tau_i$ is real. After fitting for all other parameters, we fit for the reflection mode. 
\textbf{We found that if we did not first fit and remove the polynomial terms, our reflection
fits would be dominated by the large power in the smooth structure, resulting in solutions that
did not match physical reality.
The reflection fitting is}
done for antennas with suspected reflections, the most offensive of which is seen in the 
antennas with cables of 150 meters. Because the cables were not cut at exact lengths 
(variations on order tens of centimeters), we solve for both the reflection coefficient and 
delay by performing a direct Fourier transform to a highly over resolved delay grid and 
selecting the mode where the reflection amplitude is largest. 
This fitting produces a single reflection coefficient for each antenna with nominal cable
length of 150 meters.

@@

\textbf{This effect is distinct from more fundamental wedge effects 
\citep[e.g.][]{Morales:2012, Hazelton:2013} in that it is a result of the computational limitation
of the analysis. The primary contribution to the foreground mode mixing is due to
the loss of information as each baseline samples a range of $(u,v)$ modes. On the other
hand, the effect discussed here is due to small positional offsets in the gridding kernel
and only exists after gridding.}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{1D_spectra_comparison.pdf}
\caption{
Here we compare the upper limits from the FHD$+$\eppsilon pipeline with those from the 
RTS$+$CHIPS pipeline. The magenta lines are repeated from Figure~\ref{fig:1d_ps}, and the gray lines
are the power plus two sigma thermal error bars lines generated by CHIPS. 
Due to the logarithmic binning used by CHIPS, some bins are empty, resulting in gaps in the
spectra.
CHIPS does not 
excise the coarse band harmonic lines, and so we see their strong effect where there are 
gaps in the magenta lines. 
\textbf{The RTS$+$CHIPS pipeline accesses larger $k$ modes because the data is
processed at 80~kHz, whereas the FHD$+$\eppsilon pipeline averages to 160~kHz when
gridding.}
The most notable differences are the ability of the RTS$+$CHIPS pipe
to access lower $k$ modes, while the FHD$+$\eppsilon pipe suppresses more power before and
after the first coarse band harmonic. While the different analysis methods result in slightly
different upper limits, they are generally in agreement. 
\label{fig:ps_comparison}
}
\end{figure}
